### 2014-04-29 doc/coppermtn_multilevel/tech_report/multilevel_mc.tex
adding heterogenous multilevel figures to mlmc tech note

\subsection{Heterogeneous Results}
We next solve the same linear problem as before using $M=2$ but this
time with a constant forcing of $\ve{b}=1.0$ such that:
\begin{equation}
  \ve{A}\ve{x} = \ve{b}\:.
  \label{eq:heterogeneous_problem}
\end{equation}
We will choose an initial guess of $\ve{0}$ for these calculations as
the solution is non-trivial. Studies indicate that residual norms are
controlled by the convergence of the finest level. However, adding
additional levels provides significant smoothing to the statistical
error. For example, given a fixed amount of compute time of
approximately 7 seconds, Eq~(\ref{eq:heterogeneous_problem}) was
solved using 1 level and 10,000 samples and again using 9 levels and
1,546,783 total samples ($N=1,000,000$). The 1 level calculation gave
$||\ve{r}||_\infty = 0.027$ and $||\ve{r}||_2 = 0.230$ while the 1
level calculation gave $||\ve{r}||_\infty = 0.129$ and $||\ve{r}||_2 =
0.812$. Based on the residual norm values, it would seem as though the
1 level result was superior as better convergence was achieved in the
same amount of compute time. However, the structure of the actual
solutions would indicate otherwise. Figure~\ref{fig:7_sec_1_level}
gives the 1 level result and Figure~\ref{fig:7_sec_9_level} gives the
9 level result, both plotted against an exact solution computed using
a minimal residual iteration. Clearly, the estimate produced by the 9
level computation is more representative of the exact solution than
the single level scheme due to the well-converged results provided
coarse level estimates.
\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=4in]{7_sec_1_level.pdf}
  \end{center}
  \caption{\textbf{Solution of the Poisson problem using the geometric
      multilevel Monte Carlo method with 1 level and 6.87 seconds of
      compute time.}}
  \label{fig:7_sec_1_level}
\end{figure}
\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=4in]{7_sec_9_level.pdf}
  \end{center}
  \caption{\textbf{Solution of the Poisson problem using the geometric
      multilevel Monte Carlo method with 9 levels and 6.78 seconds of
      compute time.}}
  \label{fig:7_sec_9_level}
\end{figure}
To show that the 1 level calculation will eventually converge to the
correct result, Figure~\ref{fig:converged_1_level} shows the solution
with 1,000,000 samples. This calculation required 665.7 seconds to run
and gave $||\ve{r}||_\infty = 0.0025$ and $||\ve{r}||_2 = 0.022$.
\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=4in]{converged_1_level.pdf}
  \end{center}
  \caption{\textbf{Solution of the Poisson problem using the geometric
      multilevel Monte Carlo method with 1 level and 665.7 seconds of
      compute time.}}
  \label{fig:converged_1_level}
\end{figure}

@@

>>>>

\subsection{Heterogeneous Results}
We next solve the same linear problem as before using $M=2$ but this
time with a constant forcing of $\ve{b}=1.0$ such that:
\begin{equation}
  \ve{A}\ve{x} = \ve{b}\:.
  \label{eq:heterogeneous_problem}
\end{equation}
We will choose an initial guess of $\ve{0}$ for these calculations as
the solution is non-trivial and fix $N$ at 10,000 while increasing the
number of levels. Table~\ref{tab:heterogeneous_fixed_N} gives the
results of these calculations. Here the relative error error
calculations were computed using an exact solution computed using a
minimal residual iteration converged to a residual norm tolerance of
$\sn{1}{-12}$.
\begin{table}[h!]
  \begin{center}
    \begin{tabular}{cllllllll}\hline\hline
      \multicolumn{1}{c}{\textbf{Levels}} &
      \multicolumn{1}{l}{\textbf{Samples}} &
      \multicolumn{1}{l}{\textbf{$||\ve{r}||_{\infty}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{r}||_{2}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{e}||_{\infty}$}} &
      \multicolumn{1}{l}{\textbf{$||\ve{e}||_{2}$}} &
      \multicolumn{1}{l}{\textbf{Time (s)}} &
      \multicolumn{1}{l}{\textbf{RRESFOM}} &
      \multicolumn{1}{l}{\textbf{RERRFOM}} \\
      \hline
      1 & 10,000 & 0.025 & 0.223 & 0.241 & 2.97 & 6.77 & 1.0 & 1.0 \\
      2 & 17,596 & 0.044 & 0.339 & 0.277 & 5.49 & 7.19 & 0.31 & 0.71 \\
      3 & 29,571 & 0.050 & 0.365 & 0.131 & 1.66 & 7.21 & 0.23 & 3.18 \\
      4 & 53,295 & 0.062 & 0.483 & 0.081 & 0.96 & 7.29 & 0.16 & 8.28 \\
      5 & 92,301 & 0.086 & 0.662 & 0.105 & 0.88 & 7.13 & 0.08 & 5.00 \\
      6 & 185,266 & 0.120 & 0.891 & 0.102 & 0.72 & 7.10 & 0.04 & 5.32 \\
      7 & 355,543 & 0.069 & 0.513 & 0.056 & 0.51 & 6.92 & 0.13 & 18.12 \\
      8 & 742,337 & 0.093 & 0.654 & 0.081 & 0.67 & 7.07 & 0.07 & 8.47 \\
      9 & 1,624,121 & 0.122 & 0.758 & 0.107 & 0.83 & 7.08 & 0.04 &
      4.85 \\
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Geometric Multilevel Monte Carlo results for
      $\ve{b} = 1$ with $M = 2$, $t \approx 7s$ and run times reported
      in seconds.} \textit{RRESSFOM values are the relative figure of
      merit computed using the infinity norm of the residual. RERRFORM
    values are the relative figure of merit computed using the
    infinity norm of the absolute error vector.}}
  \label{tab:heterogeneous_fixed_N}
\end{table}
Based on the residual norm values, it would seem as though the 1 level
result was superior as better convergence was achieved in the same
amount of compute time using residual norm metrics. However, the
structure of the actual solutions would indicate
otherwise. Figure~\ref{fig:heterogeneous_ml_solutions} gives the
results for with 1-9 levels all plotted against the exact
solution. Clearly, the estimate produced by the 9 level computation is
more representative of the exact solution than the single level scheme
as indicated by the error norms due to the well-converged results
provided coarse level estimates. Furthermore, the error norms also
indicate a more efficient solution scheme when multiple levels are
used showing similar relative figure of merit metrics for the error
norm computations.

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_1_level.pdf}
    \caption{\textbf{1 level.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_2_level.pdf}
    \caption{\textbf{2 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_3_level.pdf}
    \caption{\textbf{3 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_4_level.pdf}
    \caption{\textbf{4 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_5_level.pdf}
    \caption{\textbf{5 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_6_level.pdf}
    \caption{\textbf{6 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_7_level.pdf}
    \caption{\textbf{7 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_8_level.pdf}
    \caption{\textbf{8 levels.}}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{7_sec_9_level.pdf}
    \caption{\textbf{9 levels.}}
  \end{subfigure}
  \caption{\textbf{Solution of the Poisson problem using the geometric
      multilevel Monte Carlo method with 1-9 levels and approximately
      7 seconds of compute time.}}
  \label{fig:heterogenous_ml_solutions}
\end{figure}

To show that the 1 level calculation will eventually converge to the
correct result, Figure~\ref{fig:converged_1_level} shows the solution
with 1,000,000 samples. This calculation required 665.7 seconds to run
and gave $||\ve{r}||_\infty = 0.0025$ and $||\ve{r}||_2 = 0.022$.
\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=4in]{converged_1_level.pdf}
  \end{center}
  \caption{\textbf{Solution of the Poisson problem using the geometric
      multilevel Monte Carlo method with 1 level and 665.7 seconds of
      compute time.}}
  \label{fig:converged_1_level}
\end{figure}

@@

%%---------------------------------------------------------------------------%%
\section{Coarse Grid Monte Carlo Acceleration}
Based on the above results we then consider using only the coarse grid
calculations as the acceleration. Consider the following results where
a coarse grid solution has been produced and then simply prolongated
back to the fine grid rather than solving all levels as a simple
coarse approximation of the solution.
